### 2015-03-25 Kokkos_PG_Compilation.tex
Upgraded the embedded Makefile description for the new Makefile.kokkos

Kokkos consists mainly of header files. 
Only a few functions have to be compiled into object files outside of the application's source code.
Those functions are contained in \verb!.cpp! files inside the \lstinline|kokkos/core/src| directory and its subdirectories.
The files are internally protected with macros to prevent compilation if the related execution space is not enabled. 
Thus, it is not necessary to create a list of included object files specific to your compilation target.
One may simply compile all \verb!.cpp! files. 
The enabled features are controlled via macros which have to be provided in the compilation line or in the \lstinline|KokkosCore_config.h| include file. 
A list of macros can be found in Table \ref{TBL:CompileMacros}.
\begin{table}
\caption{Table of configuration Macros}
\label{TBL:CompileMacros}
\begin{small}
\begin{tabular}[t]{p{0.25\textwidth}p{0.3\textwidth}p{0.42\textwidth}}
\hline\hline
Macro & Effect & Comment \\
\hline
{\tiny KOKKOS\_HAVE\_CUDA} & Enable the CUDA execution space. & Requires a compiler capable of understanding CUDA-C. See Section \ref{S:build:CUDA}. \\
{\tiny KOKKOS\_HAVE\_OPENMP} & Enable the OpenMP execution space. & Requires the compiler to support OpenMP (e.g., \verb!-fopenmp!). \\
{\tiny KOKKOS\_HAVE\_PTHREADS} & Enable the Threads execution space. & Requires linking with libpthread.\\
{\tiny KOKKOS\_HAVE\_CXX11} & Enable internal usage of C++11 features. & The code needs to be compile with the C++11 standard. Most compilers accept the \verb!-std=c++11! flag for this.\\
{\tiny KOKKOS\_HAVE\_HWLOC} & Enable thread and memory pinning via hwloc. & Requires linking with libhwloc.\\
{\tiny KOKKOS\_DONT\_INCLUDE \_KOKKOSCORE\_CONFIG\_H} & Do not include the KokkosCore\_config.h file. & Useful when providing all macros via the compile line instead of auto-generating that file. 
This option is set for the embedded Makefile. \\
\hline\hline
\end{tabular}
\end{small}
\end{table}

@@

Kokkos provides an embedded Makefile (\lstinline|kokkos/Makefile.kokkos|) for inclusion in an application Makefile. 
This embedded Makefile generates a \lstinline|KOKKOS_INC| and a \newline
\lstinline|KOKKOS_LINK| variable which need to be appended to the compile and the link line respectively in the application Makefile. 
If you want to include dependencies (i.e. trigger a rebuild of the application object files when Kokkos files change) you can include \lstinline|KOKKOS_HEADERS| in you dependency list.
Note that you cannot compile and link at the same time!
Makefile.kokkos uses a number of variables which can be set either in the parent Makefile prior to including it, or on the command line.
These variables take the form of \lstinline|SOME_VAR=yes/no| or \lstinline|SOME_PATH=/path/to/library| and control build options such as the target architecture. 
A list can be found in Table \ref{TBL:MakefileOptions}.
Example application Makefiles can be found in the tutorial examples under \verb!kokkos/example/tutorial!.

\begin{table}
\caption{Table of Makefile options}
\label{TBL:MakefileOptions}
\begin{small}
\begin{tabular}[t]{lp{0.7\textwidth}}
\hline\hline
Option & Description \\
\hline
CUDA & Enable the CUDA execution space. \\
OMP & Enable the OpenMP execution space. \\
PTHREADS & Enable the Threads execution space.\\
MIC & Enable compilation for Intel Xeon Phi. \\
CXX11 & Enable internal usage of C++11 features.\\
HWLOC & Enable hwloc usage.\\
LIBRT & Enable librt usage for the \lstinline|Kokkos::Impl::Timer| class.\\
CUDA\_PATH & Set the path to the CUDA toolkit directory.\\
HWLOC\_PATH & Set the path to the hwloc library.\\
\hline\hline
\end{tabular}
\end{small}
\end{table}

>>>>

Kokkos consists mainly of header files. 
Only a few functions have to be compiled into object files outside of the application's source code.
Those functions are contained in \verb!.cpp! files inside the \lstinline|kokkos/core/src| directory and its subdirectories.
The files are internally protected with macros to prevent compilation if the related execution space is not enabled. 
Thus, it is not necessary to create a list of included object files specific to your compilation target.
One may simply compile all \verb!.cpp! files. 
The enabled features are controlled via macros which have to be provided in the compilation line or in the \lstinline|KokkosCore_config.h| include file. 
A list of macros can be found in Table \ref{TBL:CompileMacros}.
\begin{table}
\caption{Table of configuration Macros}
\label{TBL:CompileMacros}
\begin{small}
\begin{tabular}[t]{p{0.25\textwidth}p{0.3\textwidth}p{0.42\textwidth}}
\hline\hline
Macro & Effect & Comment \\
\hline
{\tiny KOKKOS\_HAVE\_CUDA} & Enable the CUDA execution space. & Requires a compiler capable of understanding CUDA-C. See Section \ref{S:build:CUDA}. \\
{\tiny KOKKOS\_HAVE\_OPENMP} & Enable the OpenMP execution space. & Requires the compiler to support OpenMP (e.g., \verb!-fopenmp!). \\
{\tiny KOKKOS\_HAVE\_PTHREADS} & Enable the Threads execution space. & Requires linking with libpthread.\\
{\tiny KOKKOS\_HAVE\_Serial} & Enable the Serial execution space. & \\
{\tiny KOKKOS\_HAVE\_CXX11} & Enable internal usage of C++11 features. & The code needs to be compile with the C++11 standard. Most compilers accept the \verb!-std=c++11! flag for this.\\
{\tiny KOKKOS\_HAVE\_HWLOC} & Enable thread and memory pinning via hwloc. & Requires linking with libhwloc.\\
\hline\hline
\end{tabular}
\end{small}
\end{table}

@@

The base of the build system is the Makefile.kokkos, which is designed to be included by application Makefiles.
It contains logic to (re)generate the KokkosCore\_config.h file if necessary, build the Kokkos library, and provide updated compiler and linker flags.

The system can digest a number of variables which are used to configure Kokkos settings. 
Generally the variables are parsed for Keywords. 
This allows for multiple options being given for each variable. 
The separator doesn't matter as long as it doesn't interact with the Make system.
A list of variables, their meaning and options is given in table \ref{TBL:MakefileOptions}.

\begin{table}
\caption{Variables for the embedded Makefile}
\label{TBL:MakefileOptions}
\begin{small}
\begin{tabular}[t]{p{0.3\textwidth}p{0.65\textwidth}}
\hline\hline
Variable & Description \\
\hline
   KOKKOS\_PATH (IN): & Path to the Kokkos root or install directory. 
                   One can either build against an existing install 
                   of Kokkos or use its source directly for an 
                   embedded build. In the former case the "Input variables"
                   are set inside the embedded Makefile.kokkos and it 
                   is not valid to set them differently in the including
                   Makefile.\\

\hline
   CUDA\_PATH (IN): & Path to the Cuda toolkit root directory.\\

\hline
   HWLOC\_PATH (IN): & Path to the hardware locality library. \\

\hline
   KOKKOS\_DEVICES (IN): & What Execution and Memory Spaces should be enabled. \\
      \hspace{0.5cm}Options: & OpenMP,Serial,Pthreads,Cuda \\
      \hspace{0.5cm}Default: & OpenMP \\

\hline    
    KOKKOS\_ARCH (IN): & What backend Architecture to build for. \\
      \hspace{0.5cm}Options: & KNC,SNB,HSW,Kepler,Kepler30,Kepler35,Kepler37,Maxwell,
               Maxwell50,ARMv8,BGQ,Power7,Power8 \\
      \hspace{0.5cm}Default: & None -- this means no particular architecture flags
               are set. \\

\hline
    KOKKOS\_DEBUG (IN): & Enable debugging. \\
      \hspace{0.5cm}Options: & yes,no \\
      \hspace{0.5cm}Default: & no \\

\hline
    KOKKOS\_CXX\_STANDARD (IN): & Set the C++ standard to be used. \\
      \hspace{0.5cm}Options: & c++11 \\
      \hspace{0.5cm}Default: & c++11 \\

\hline\hline

    KOKKOS\_CPPFLAGS (OUT): & Pre processor flags (include directories and defines). 
                              Add this to applications compiler and preprocessor flags. \\

\hline
    KOKKOS\_CXXFLAGS (OUT): & Compiler flags. Add this to the applications compiler flags. \\

\hline
    KOKKOS\_LDFLAGS (OUT): & Linker flags. Add this to the applications linker flags. \\

\hline
    KOKKOS\_LIBS (OUT): & Libraries required by Kokkos. Add this to the link line after the linker flags. \\

\hline
    KOKKOS\_CPP\_DEPENDS (OUT): & Dependencies for compilation units which include any Kokkos header files.
                                  Add this as a dependency to compilation targets including any Kokkos code. \\

\hline
    KOKKOS\_LINK\_DEPENDS (OUT): & Dependencies of an application linking in the Kokkos library.
                                   Add this to the dependency list of link targets. \\

\hline\hline
    CXXFLAGS (IN): & User provided compiler flags which will be used to compile the Kokkos library. \\

\hline
    CXX (IN): & The compiler used to compile the Kokkos library. \\
\hline\hline
\end{tabular}
\end{small}
\end{table}

    A word of caution on where to include the Makefile.kokkos:
    since the embedded Makefiles defines targets it is usually
    better to include it after the first application target
    has been defined. Since that target can't use the flags
    from the embedded Makefiles it should be a meta target:

\begin{lstlisting}    
    CXX=g++

    default: main

    include Makefile.kokkos

    main: $(KOKKOS_LINK_DEPENDS) $(KOKKOS_CPP_DEPENDS) main.cpp
        $(CXX) $(KOKKOS_CPPFLAGS) $(KOKKOS_CXXFLAGS) \
        $(KOKKOS_LDFLAGS) $(KOKKOS_LIBS) main.cpp -o main
\end{lstlisting}

More example application Makefiles can be found in the tutorial examples under \verb!kokkos/example/tutorial!.

### 2015-03-25 Kokkos_PG_Views.tex
Fixed missing character
Upgraded the embedded Makefile description for the new Makefile.kokkos

C++ lets users construct data types that may ``look like'' numbers in
terms of syntax, but do arbitrarily complicated things inside.  Some
of those things may not be thread safe, like unprotected updates to
global state.  Others may perform badly in parallel, like fine-grained
dynamic memory allocation.  Therefore it is strongly advised to use 
only simple data types inside Kokkos Views.

Users may always construct a View whose entries are 
\begin{itemize}
\item built-in data types (``plain old data''), like \texttt{int} or
  \texttt{double}, or
\item structs of built-in data types.
\end{itemize}

@@

You may access an entry of a View using parentheses enclosing a
comma-delimited list of integer indices.  This looks just like a
Fortran multidimensional array access.  For example:
\begin{lstlisting}
const size_t N = ...;
View<double*[3][4]> a ("some label", N);
// KOKKOS_LAMBDA macro includes capture-by-value specifier [=].
parallel_for (N, KOKKOS_LAMBDA (const ptrdiff_t i) {
    const size_t j = ...;
    const size_t k = ...;
    const double a_ijk = a(i,j,k);
    /* rest of the loop body */
  });
\end{lstlisting}
Note how in the above example, we only access the View's entries in a
parallel loop body.  In general, you may only access a View's entries
in an execution space which is allowed to access that View's memory
space.  For example, if the default execution space is \lstinline!Cuda!, 
a View for which no specific Memory Space was fiven may not be accessed 
in host code \footnote{An exemption is if you specified for CUDA compilation
that the default memory space is CudaUVMSpace, which can be accessed from 
the host.}
Furthermore, access costs (e.g., latency and
bandwidth) may vary, depending on the View's ``native'' memory and
execution spaces, and the execution space from which you access it.
CUDA UVM may work, but it may also be slow, depending on your access
pattern and performance requirements.  Thus, best practice is to
access the View only in a Kokkos parallel for, reduce, or scan, using
the same execution space as the View.  This also ensures that access
to the View's entries respect first-touch allocation.  The first
(leftmost) dimension of the View is the \emph{parallel dimension},
over which it is most efficient to do parallel array access if the 
default memory layout is used (e.g. if no specific memory layout is 
specified).

@@

Const Views often allow the compiler to optimize more agressively by allowing it
to reason about possible write conflicts and data aliasing. For example in a vector
update \lstinline|a(i+1)+=b(i) with skewed indexing it is safe to vectorize if
\lstinline|b| is a View of const data.

>>>>

C++ lets users construct data types that may ``look like'' numbers in
terms of syntax, but do arbitrarily complicated things inside.  Some
of those things may not be thread safe, like unprotected updates to
global state.  Others may perform badly in parallel, like fine-grained
dynamic memory allocation.  Therefore it is strongly advised to use 
only simple data types inside Kokkos Views.
Users may always construct a View whose entries are 
\begin{itemize}
\item built-in data types (``plain old data''), like \texttt{int} or
  \texttt{double}, or
\item structs of built-in data types.
\end{itemize}

@@

You may access an entry of a View using parentheses enclosing a
comma-delimited list of integer indices.  This looks just like a
Fortran multidimensional array access.  For example:
\begin{lstlisting}
const size_t N = ...;
View<double*[3][4]> a ("some label", N);
// KOKKOS_LAMBDA macro includes capture-by-value specifier [=].
parallel_for (N, KOKKOS_LAMBDA (const ptrdiff_t i) {
    const size_t j = ...;
    const size_t k = ...;
    const double a_ijk = a(i,j,k);
    /* rest of the loop body */
  });
\end{lstlisting}
Note how in the above example, we only access the View's entries in a
parallel loop body.  In general, you may only access a View's entries
in an execution space which is allowed to access that View's memory
space.  For example, if the default execution space is \lstinline!Cuda!, 
a View for which no specific Memory Space was given may not be accessed 
in host code \footnote{An exemption is if you specified for CUDA compilation
that the default memory space is CudaUVMSpace, which can be accessed from 
the host.}.
Furthermore, access costs (e.g., latency and
bandwidth) may vary, depending on the View's ``native'' memory and
execution spaces, and the execution space from which you access it.
CUDA UVM may work, but it may also be slow, depending on your access
pattern and performance requirements.  Thus, best practice is to
access the View only in a Kokkos parallel for, reduce, or scan, using
the same execution space as the View.  This also ensures that access
to the View's entries respect first-touch allocation.  The first
(leftmost) dimension of the View is the \emph{parallel dimension},
over which it is most efficient to do parallel array access if the 
default memory layout is used (e.g. if no specific memory layout is 
specified).

@@

Const Views often allow the compiler to optimize more agressively by allowing it
to reason about possible write conflicts and data aliasing. For example in a vector
update \lstinline|a(i+1)+=b(i)| with skewed indexing it is safe to vectorize if
\lstinline|b| is a View of const data.
